Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-27 main.tex
def complete label theory & prop

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)

Let $\Semiring$ be a commutative semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.

@@

$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,

@@

\wei_{10}(q, a, \epsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)

@@

\wei_{01}(q, \epsilon, b, q') & = & \psi(b) & 

@@

\weight_T(q, \epsilon, \epsilon, q)  & = \final(q)\\
\weight_T(q, \epsilon, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, au, \epsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, \epsilon, q')\nonumber\\
\weight_T(q, \epsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', \epsilon, v, q')\nonumber\\

@@

    \wei_{10}(q, a, \epsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\

@@

    \wei_{01}(q, \epsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\

@@

When $a = \epsilon$ (resp. $b = \epsilon$), the current symbol 

@@

%whereas considering the current symbol may be useful to compute a transition weight.
%(even when it does not change the head's position, like with $\epsilon$-transitions).
%
%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
%are missing in the definition of $\weight_T$.
%It means that $T$ must avoid configurations where it reached the end of 
%the output word and not of the input one, or vice-versa.
%This can be done by using $\wei_{10}$ and $\wei_{01}$
%before reaching the end of word, and using a special state for this purpose.

The expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei_{10}(q, a, \epsilon, r)$ is the cost of 
the deletion of the symbol~$a \in \Sigma$ in~$s$, 
$\wei{01}(q, \epsilon, b, r)$ is the cost 
of the insertion of~$b \in \Delta$ in $t$, 
and $\wei_{11}(q, a, b, r)$ is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

@@

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \epsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \epsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and 
$\init(q_1) = \final(q_1) = \zero$. 
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and ${\wei_1}$ is a transition functions 
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
%      
\noindent
As above in the case of \SWT, 
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
The computation of $A$ on words $s \in \Sigma^*$
is defined with an intermediate function $\weight_A$, 
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
%
\begin{align}
\weight_A(q, \epsilon, q) & = \final(q)\\
\weight_A(q, \epsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to 
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_A(q, s, q') 
\label{eq:SWA-value}
\end{equation}

%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
and $s \in \Sigma^+$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
%where the component functions of
%be the synthesized form of 
%$Q \times (\Sigma \cup \{ \epsilon \}) \times (\Delta \cup \{ \epsilon \}) \times Q \to \Semiring$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$). %$n = |s|$ be the length of $s$.

@@

Every non-null transition of $A_{T, s}$ will
simulate a sequence of transitions of $T$ performing the following steps:
advance in the input word while staying immobile in the output word, 
and then make one step in the output word (and advance in the input word or not).
%
The first steps in the sequence correspond to $\epsilon$-transitions of automata, 
and their total weight is computed by the following intermediate function 
$\wei'_0 : Q' \times Q' \to \Semiring$, 
defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
by:
\[
\begin{array}{rcl}
\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
\quad \mathrm{if~} 1 \leq i \leq n+1,\\
%
\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
%
\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
\displaystyle
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
           \scriptstyle q_0 = q, q_k = q'
           \end{array}} 
\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
\quad \begin{array}[t]{l}
      \mathrm{if~} 1 \leq i < n,
      \mathrm{~and~} 1 \leq k \leq n-i,\\
      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
      \end{array}
\end{array}
\]
%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
%
The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
(Section~\ref{section:symbols}). 
% closed under partial applications 
The sum and product in its definition are finite, 
we shall see below how to compute the first sum in polynomial time.
%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
%of length $k+1$ in the sum.
%Indeed, it amounts to search short

@@

Formally, the transitions of the automaton~$A$ are defined
in term of %a weight value computed by 
an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
Hence, $\weight_A$ is a function from 
$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
%
\begin{equation}
\begin{array}[t]{rcl}
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & = & 
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\\
\end{array}
\label{eq:SWVPA-weight}
\end{equation}
%
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
and $\gamma$ is the rest of stack.

@@

% A \emph{regular tree grammar} over $\Omega$ 
% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
% $N$ is a finite set of non-terminal symbols denoted $q$..., 
% $q_\mathsf{i} \in N$ is the starting non-terminal, 
% $R$ is a finite set of production rules of the form
% $q_0 \to a(q_1\ldots q_k)$ where 
% $q_0, q_1, \ldots, q_k \in N$
% $a \in \Omega_k$.
% A tree $t \in \T(\Omega)$ is in the language of $G$ 
% if it can be generated from $q_\mathsf{i}$ by 
% non terminal replacement following the rules of $R$.
%
\begin{definition}  \label{def:SWTA}
A \emph{symbolic-weighted tree automaton} (\SWTA)
over $\Omega$, $\Semiring$, and $\bar\Phi$
is a triplet $A = \< Q, \init, \bar{\wei} >$ where
$Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
and $\bar{\wei}$ is a tuplet of transition functions containing
%$\wei_\epsilon$ from $Q \times Q$ into $\Semiring$, and, 
$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
for each $k \in [0..M]$. 
\end{definition}
%
%Like in Section~\ref{sec:SWAdef}, 
We define from $\bar{\wei}$ 
a transition function~$\wei$, %from $\bar\wei$
from 
$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
into~$\Semiring$ by: %also called $\wei$ for simplicity, 
%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
\[
\begin{array}{rcll}
%\wei(q_0, \epsilon, q_1) & = &  \wei_\epsilon(q_0, q_1),\\ %\phi_\epsilon\\
\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
\end{array}      
\]
Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
of a regular tree grammar~\cite{tata}, 
that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
%in a step of tree building.
%
Such a grammar computes the weights of the derivation trees 
of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.

>>>>

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)
\noindent 
Let $\Semiring$ be a commutative semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be a label theory over $\Semiring$.

@@

$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,

@@

\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)

@@

\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 

@@

\weight_T(q, \varepsilon, \varepsilon, q)  & = \final(q)\\
\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, au, \varepsilon, q') & = \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, \varepsilon, q')\nonumber\\
\weight_T(q, \varepsilon, bv, q') & = \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', \varepsilon, v, q')\nonumber\\

@@

    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, bv, q')\nonumber\\

@@

    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', au, v, q')\nonumber\\

@@

When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 

@@

%whereas considering the current symbol may be useful to compute a transition weight.
%(even when it does not change the head's position, like with $\varepsilon$-transitions).
%
%The cases $\weight_T(q, au, \varepsilon, q')$ and $\weight_T(q, \varepsilon, bv, q')$
%are missing in the definition of $\weight_T$.
%It means that $T$ must avoid configurations where it reached the end of 
%the output word and not of the input one, or vice-versa.
%This can be done by using $\wei_{10}$ and $\wei_{01}$
%before reaching the end of word, and using a special state for this purpose.

The expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei_{10}(q, a, \varepsilon, r)$ is the cost of 
the deletion of the symbol~$a \in \Sigma$ in~$s$, 
$\wei{01}(q, \varepsilon, b, r)$ is the cost 
of the insertion of~$b \in \Delta$ in $t$, 
and $\wei_{11}(q, a, b, r)$ is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

@@

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and 
$\init(q_1) = \final(q_1) = \zero$. 
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and 
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$, 
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$, %\Phi_\Sigma$, 
respectively $\mathsf{out} : Q \to \Semiring$  %\Phi_\Sigma,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and ${\wei_1}$ is a transition functions 
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
%      
\noindent
As above in the case of \SWT, 
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
The computation of $A$ on words $s \in \Sigma^*$
is defined with an intermediate function $\weight_A$, 
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
%
\begin{align}
\weight_A(q, \varepsilon, q) & = \final(q)\\
\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to 
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q) 
\mathop{\otimes} \weight_A(q, s, q') 
\label{eq:SWA-value}
\end{equation}

%
%When $\wei_\varepsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\varepsilon$-transitions}.

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
and $s \in \Sigma^+$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^+$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
with $\bar{\wei} = \< \wei_{10}, \wei_{01}, \wei_{11}>$,
%where the component functions of
%be the synthesized form of 
%$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$ ($n \geq 1$). %$n = |s|$ be the length of $s$.

@@

Every non-null transition of $A_{T, s}$ will
simulate a sequence of transitions of $T$ performing the following steps:
advance in the input word while staying immobile in the output word, 
and then make one step in the output word (and advance in the input word or not).
%
The first steps in the sequence correspond to $\varepsilon$-transitions of automata, 
and their total weight is computed by the following intermediate function 
$\wei'_0 : Q' \times Q' \to \Semiring$, 
defined for all $q, q' \in Q$: % all $1 \leq i < n$ and $1 \leq k \leq n-i$, 
by:
\[
\begin{array}{rcl}
\wei'_0\bigl(\<i, q>, \< i, q>\bigr) & = & \one 
\quad \mathrm{if~} 1 \leq i \leq n+1,\\
%
\wei'_0\bigl(\<i, q>, \< i, q'>\bigr) & = & \zero 
\quad \mathrm{if~} 1 \leq i \leq n+1 \mathrm{~and~} q \neq q',\\
%
\wei'_0\bigl(\<i, q>, \< i+k, q'>\bigr) & = & 
\displaystyle
\bigoplus_{\begin{array}{c}
           \scriptstyle q_0,\ldots, q_k \in Q\\[-2pt]
           \scriptstyle q_0 = q, q_k = q'
           \end{array}} 
\bigotimes_{j=0}^{k-1} \phi_{j, j+1}(s_{i+j})
\quad \begin{array}[t]{l}
      \mathrm{if~} 1 \leq i < n,
      \mathrm{~and~} 1 \leq k \leq n-i,\\
      \mathrm{where~} \phi_{j, j+1} = \wei_{10}(q_j, q_{j+1})
      \end{array}
\end{array}
\]
%where $\bigl[\wei_{10}(q_j, q_{j+1})\bigr]_{s_{i+j}}$ 
%is the partial application $\eta_{s_{i+j}} \in \Phi_{\Delta}$
%for $\eta = \wei_{10}(q_j, q_{j+1}) \in \Phi_{\Sigma, \Delta}$ ($s_{i+j} \in \Sigma$).
%
The function $\wei'_0$ is defined thanks to the closure properties of the label theory $\bar\Phi$
(Section~\ref{section:symbols}). 
% closed under partial applications 
The sum and product in its definition are finite, 
we shall see below how to compute the first sum in polynomial time.
%The effective computation of $\wei'_0$ does not require enumerating all the state sequences 
%of length $k+1$ in the sum.
%Indeed, it amounts to search short

@@

Formally, the transitions of the automaton~$A$ are defined
in term of %a weight value computed by 
an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
In the case of a pushdown automaton, a configuration is composed 
of a state $q \in Q$ and a stack content $\gamma \in \Gamma^*$, where $\Gamma = \Omegac \times P$.
Hence, $\weight_A$ is a function from 
$[Q \times \Gamma^*] \times \Omega^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
%
\begin{align}
%\begin{array}[t]{rcl}
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, a\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
\weight_A\bigl(\config{q}{\bot}, a\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot\gamma}, {\call{c}'}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-2pt]
                          \scriptstyle p \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}'}\, p' \cdot {\call{c}}\, p\cdot \gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weiec(q, {\call{c}}, p, q'') 
 \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{{\call{c}}\, p\cdot \gamma}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} 
 \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
 \otimes \weight_A\bigl(\config{q''}{\gamma}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
 \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)
%\end{array}
\label{eq:SWVPA-weight}
\end{align}
%
where $\bot$ denotes the empty stack and ${\call{c}}\, p\cdot \gamma$ 
denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
and $\gamma$ is the rest of stack.

@@

% A \emph{regular tree grammar} over $\Omega$ 
% is a triplet $G = \< N, q_\mathsf{i}, R>$ where
% $N$ is a finite set of non-terminal symbols denoted $q$..., 
% $q_\mathsf{i} \in N$ is the starting non-terminal, 
% $R$ is a finite set of production rules of the form
% $q_0 \to a(q_1\ldots q_k)$ where 
% $q_0, q_1, \ldots, q_k \in N$
% $a \in \Omega_k$.
% A tree $t \in \T(\Omega)$ is in the language of $G$ 
% if it can be generated from $q_\mathsf{i}$ by 
% non terminal replacement following the rules of $R$.
%
\begin{definition}  \label{def:SWTA}
A \emph{symbolic-weighted tree automaton} (\SWTA)
over $\Omega$, $\Semiring$, and $\bar\Phi$
is a triplet $A = \< Q, \init, \bar{\wei} >$ where
$Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ is the starting weight function, 
and $\bar{\wei}$ is a tuplet of transition functions containing
%$\wei_\varepsilon$ from $Q \times Q$ into $\Semiring$, and, 
$\wei_{k}: Q \times Q^{k} \to \Phi_{\Omega_k}$
for each $k \in [0..M]$. 
\end{definition}
%
%Like in Section~\ref{sec:SWAdef}, 
We define from $\bar{\wei}$ 
a transition function~$\wei$, %from $\bar\wei$
from 
$Q \times \Omega \times \bigcup_{k=0}^{M} Q^k$
into~$\Semiring$ by: %also called $\wei$ for simplicity, 
%such that, for all $q, q' \in Q$, $a \in \Sigma$, and $b \in \Delta$, 
\[
\begin{array}{rcll}
%\wei(q_0, \varepsilon, q_1) & = &  \wei_\varepsilon(q_0, q_1),\\ %\phi_\varepsilon\\
\wei(q_0, a, q_1 \ldots q_k) & = & \phi_{\Omega, k}(a) &
\quad\mathrm{where~} \phi_{k} = \wei_{k}(q_0, q_1\ldots q_k).\\
\end{array}      
\]
Intuitively, $\wei(q_0, a, q_1 \ldots q_k)$ can be seen as
the weight of a production rule $q_0 \to a(q_1, \ldots, q_k)$ 
of a regular tree grammar~\cite{tata}, 
that replaces the non-terminal symbol $q_0$ by $a(q_1, \ldots, q_k)$. 
%in a step of tree building.
%
Such a grammar computes the weights of the derivation trees 
of the Context-Free grammar obtained by forgetting the labeling symbols of $\Omega_{>0}$.
